VM-Centric Snapshot Deduplication for Cloud Data Backup

### 2015-03-07 snapshot_system/mass_arch.tex
prepare for mass conference
re

\section{Prototype Implementation}
\label{sect:architecture}

Our prototype system runs on a cluster of Linux machines with Xen-based VMs 
and the QFS~\cite{michael2013} distributed file system. 
%to %manage  the physical disk storage. 
All data needed for the backup service including snapshot data and metadata
%and snapshot data for backup purposes, 
resides in this distributed file system. 
One physical node hosts tens of VMs, each of which accesses its virtual machine disk image through the
virtual block device driver (called TapDisk\cite{Warfield2005} in Xen).

\subsection{Per Node Software Components} 
%\begin{figure}[htbp]
\begin{figure}[htbp]
%\vspace{1em}
    \centering
    %\includegraphics[width=2.75in]{images/socc_arch_cluster}
    \includegraphics[width=2.8in]{images/socc_arch_cluster}
%\vspace{1em}
    \caption{Data flow during snapshot backup}
    \label{fig:arch_vm}
\end{figure}
As depicted in Figure~\ref{fig:arch_vm}, 
there are four key service components running on each cluster
node  for supporting backup and deduplication: 
1) a virtual block device driver, 2) a snapshot deduplication agent,
3) a snapshot store client to store  and access snapshot data,
and 4)  a PDS client to support PDS metadata access.

@@

%\begin{figure}[htbp]
\begin{figure}[htbp]
%\vspace{1em}
  \centering
%  \includegraphics[width=2.75in]{images/sstore_arch}
  \includegraphics[width=2.8in]{images/sstore_arch}
%\vspace{1em}
  \centering
  \caption{VM snapshot store data structures}
  \label{fig:as_arch}
\end{figure}

>>>>

\section{System Design and Implementation}
\label{sect:architecture}

We describe the design and implementation of a VM-centric backup system that
runs on a cluster of Linux machines with Xen-based VMs 
and the QFS~\cite{michael2013} distributed file system. 
%to %manage  the physical disk storage. 
All data needed for the backup service including snapshot data and metadata
%and snapshot data for backup purposes, 
resides in this distributed file system. 
One physical node hosts tens of VMs, each of which accesses its virtual machine disk image through the
virtual block device driver (called TapDisk\cite{Warfield2005} in Xen).

\subsection{Per Node Software Components} 
%\begin{figure}[htbp]
\begin{figure}[htbp]
%\vspace{1em}
    \centering
    %\includegraphics[width=2.75in]{images/socc_arch_cluster}
    \includegraphics[width=2.7in]{images/socc_arch_cluster}
%\vspace{1em}
    \caption{Data flow during snapshot backup}
    \label{fig:arch_vm}
\end{figure}
As depicted in Figure~\ref{fig:arch_vm}, 
there are four key service components running on each cluster
node  for supporting backup and deduplication: 
1) a virtual block device driver, 2) a snapshot deduplication agent,
3) a snapshot store client to store  and access snapshot data,
and 4)  a PDS client to support PDS metadata access.

@@

%\begin{figure}[htbp]
\begin{figure}[htbp]
%\vspace{1em}
  \centering
%  \includegraphics[width=2.75in]{images/sstore_arch}
  \includegraphics[width=2.7in]{images/sstore_arch}
%\vspace{1em}
  \centering
  \caption{VM snapshot store data structures}
  \label{fig:as_arch}
\end{figure}

### 2015-03-07 snapshot_system/mass_background.tex
re

%Another disadvantage of cross-VM sharing is that it complicates
Snapshot deletion 
%which is the garbage collection in VM environment
occurs frequently because snapshots expire regularly
and reference counting is required to identify chunks that are no loner used.
%Several well developed garbage collection methods do not apply to the
%converged virtualized architecture due to various environment and resource limitations:
Grouped mark-and-sweep~\cite{Guo2011} is effective for deletion by tracking
the dependencies from data containers to metadata groups, but still it carries
a significant cost in a cloud cluster. This is because any data chunk can still be
used by a large number of VMs, and there are multiple rounds of extensive I/O involved
to scan through the meta data of snapshots to check if a chunk is referenced or not.
As the number of machines increases, the I/O cost becomes more expensive.
% and  we show  such that only a subset of
%metadata needs to be scanned in order to reclaim the garbage in any given data container.
%However, data in VM environment tend to be highly dedupable which means
%such dependency could be extensive, therefore scanning a large subset of metadata still carries a significant cost,
% especially in a distributed setting.
Perfect hashing~\cite{Fabiano2013} provides a memory efficient way to represent a large set of
hashes in a compact form, however, it still requires the scanning of all data hashes before any 
garbage collection could begin, which limits its application to an offline batch process. 
In addition, it requires a centralized location
to store the hashing result which is still too big to be held in a converged architecture.

\comments{
Our previous work on parallel deduplication~\cite{zhangusenix13} is a low-cost
solution, but it is not an inline solution and th backup for all VMs needs to be conducted
around the same time. 
It does not address  snapshot deletion and file system block packaging for fault tolerance. 
}

Our previous work on parallel deduplication~\cite{zhangusenix13,WeiZhangIEEE} 
is not a comprehensive VM-centric solution and does not address  snapshot deletion and file system block 
packaging for fault tolerance.  The work in ~\cite{zhangusenix13} is a low-cost
solution, but requires all backups for each day to occur at the same time and is less flexible than the approach presented here.

>>>>

%Another disadvantage of cross-VM sharing is that it complicates
\comments{
Snapshot deletion 
%which is the garbage collection in VM environment
occurs frequently because snapshots expire regularly
and reference counting is required to identify chunks that are no loner used.
%Several well developed garbage collection methods do not apply to the
%converged virtualized architecture due to various environment and resource limitations:
Grouped mark-and-sweep~\cite{Guo2011} is effective for deletion by tracking
the dependencies from data containers to metadata groups, but still it carries
a significant cost in a cloud cluster. This is because any data chunk can still be
used by a large number of VMs, and there are multiple rounds of extensive I/O involved
to scan through the meta data of snapshots to check if a chunk is referenced or not.
As the number of machines increases, the I/O cost becomes more expensive.
% and  we show  such that only a subset of
%metadata needs to be scanned in order to reclaim the garbage in any given data container.
%However, data in VM environment tend to be highly dedupable which means
%such dependency could be extensive, therefore scanning a large subset of metadata still carries a significant cost,
% especially in a distributed setting.
Perfect hashing~\cite{Fabiano2013} provides a memory efficient way to represent a large set of
hashes in a compact form, however, it still requires the scanning of all data hashes before any 
garbage collection could begin, which limits its application to an offline batch process. 
In addition, it requires a centralized location
to store the hashing result which is still too big to be held in a converged architecture.

Our previous work on parallel deduplication~\cite{WeiZhangIEEE}
demonstrates the simulation results for  
is not a comprehensive VM-centric solution and does not address  snapshot deletion and file system block 
packaging for fault tolerance.

Our previous work on popularity distribution of  deduplication~\cite{WeiZhangIEEE}
demonstrates the simulation results for  
The work in ~\cite{zhangusenix13} is a low-cost
solution, but requires all backups for each day to occur at the same time and is less flexible than the approach presented here.
}

\comments{
Our previous work on parallel deduplication~\cite{zhangusenix13} is a low-cost
solution, but it is not an inline solution and th backup for all VMs needs to be conducted
around the same time. 
It does not address  snapshot deletion and file system block packaging for fault tolerance.

}

### 2015-03-07 snapshot_system/mass_dedup.tex
prepare for mass conference
re

With the considerations discussed in the previous section, we propose a
VM-centric approach (called VC)
for a co-located backup service that has a resource usage profile
suitable for use with converged cloud architectures.
This compares the traditional deduplication approach {\em VM-oblivious} (VO)
which manages duplicate data chunks without consideration of VMs.
Section~\ref{sect:vc-strategies} discusses  our key ideas and design for duplicate detection.
Section~\ref{sect:delete} presents  a simplified snapshot deletion with  the VM-centric approach.

@@

We describe our overall deduplication approach as consisting of three
complementary strategies: VM-local
duplicate search, global deduplication using popular chunks, and 
VM-centric file system block management. 
\begin{itemize}
\item 
\textbf{Cross-VM global deduplication using popular chunks} --
We separate the deduplication within a VM and cross VMs
and simplify the cross-VM deduplication while maximizing the inter-VM deduplication efficiency  as much as possible.
This is because global deduplication that detects the appearance of a chunk 
in any VM requires a substantial resource for fingerprint comparison.
To simplify cross-VM deduplication, we restrict the search scope of deduplication within the top $k$ 
most popular items. This popular data set is called the \textbf{PDS}. 
%This step accomplishes the canonical global fingerprint lookup using a popular fingerprint index.
 \begin{figure}
 \centering
  \includegraphics[width=3in]{figures/zipf_count_rank.pdf}
 \caption{Duplicate frequency versus  chunk ranking in a log scale after local deduplication.}
 \label{fig:Datazipf}
 \end{figure}

% {\it [Need to find a place to put these numbers in: Total number of chunks 
% in 350 snapshots: 1,546,635,485. 
% Total number of chunks after localized dedup: 283,121,924. Total number of unique chunks: 87,692,682.]}
What  motivates us to use PDS
is that for those chunks that appear in different VMs, the top $k$ most popular items
dominate the appearance distribution~\cite{WeiZhangIEEE}.  
Figure~\ref{fig:Datazipf} shows the distribution of chunk popularity for a data trace 
from Alibaba with 2500 VMs discussed in Section ~\ref{sect:evaluation}.
We define chunk popularity as the number of unique copies of the chunk in the data-store,
i.e., the number of copies of the chunk after local deduplication.
%the distribution of chunk popularity~\cite{WeiZhangIEEE} 
%based on a production data trace from Alibaba with over 2500 VMs.
The distribution from this figure is Zipf-like. 
Let  $\sigma$ denote as the percentage of unique data chunks belonging to PDS and 
from the evaluation in Section~\ref{sect:evaluation}, we find that
$\sigma$ with about  2\% can deliver a fairly competitive deduplication efficiency.
%$\sigma$ with about  range of 2 to 4\% can deliver a fairly competitive deduplication efficiency.
%For example,
%with top 2\% of the most popular chunks stored in PDS, up-to 40\% of the raw backup 
%data can be eliminated by fingerprint lookup in PDS index.

@@

\comments{
We have not adopted the index sampling in \cite{Guo2011} because its extension for
for a distributed architecture is difficult.  To use a distributed memory version of the sampled index, every deduplication 
request may access a remote machine for index lookup and the overall overhead of access latency for all requests
can be significant.
}
\item 
\textbf{VM-specific duplicate search optimization} --
While cross-VM deduplication is simplified, we intend to optimize the VM-specific deduplication 
under a reasonable memory consumption
 to make up the loss of deduplication opportunities due to the cross-VM popularity constraint.
%Fingerprint lookup within a VM does not require a signficant  amount of memory resource.
We start with the standard version-based detection~\cite{Clements2009,Vrable2009}
to identify changed content with dirty bits in a coarse grain segment level.
%In our implementation 
%the segment size is 2MB
%and the device driver is extended to support tracking changed segments using a dirty bitmap. 
The reason to choose a coarse grain segment level is that 
since every write for a segment will touch a dirty bit, the device driver maintains dirty bits in 
memory and cannot afford a small segment size.
It should be noted that dirty bit tracking is supported or can be easily implemented in 
major virtualization solution vendors. 
%% this method is so widely-used and easy to understand, everyone include Pure Storage has it, I feel no need to explain further.
%{
%For example,
%the VMWare hypervisor has an API to let external backup applications know 
%the changed areas since last backup. 
%The Microsoft SDK provides an API that allows external applications to monitor 
%the VM's I/O traffic and implement such changed block tracking feature.
%}

@@

%The system allows all machines conduct backups in parallel, and each machine
%conducts the backup of one VM at a time, and thus only requires a write buffer for one VM.
\end{itemize}
The underlying distributed  file system has  a fixed replication degree for 
all file system blocks~\cite{googlefs03,hdfs10}.
We add extra replicas for file blocks containing PDS chunks  
and this brings additional  failure protection for these chunks shared by many VMs.
Appendix~\ref{sect:fault} provides a detailed analysis of snapshot availability
under this VC scheme and compares it with VO.

>>>>

With the considerations discussed in the previous section, we propose a
VM-centric approach (called VC)
for a co-located backup service that has a resource usage profile
suitable for use with converged cloud architectures.
This compares the traditional deduplication approach {\em VM-oblivious} (VO)
which manages duplicate data chunks without consideration of VMs.
Section~\ref{sect:vc-strategies} discusses  our key ideas and design for duplicate detection.
%Section~\ref{sect:delete} presents  a simplified snapshot deletion with  the VM-centric approach.

@@

%We describe our overall deduplication approach as consisting of three
%complementary strategies: VM-local
%duplicate search, 
%and %VM-centric file system block management. 
%\textbf{Cross-VM global deduplication using popular chunks} --
%global deduplication using popular chunks, 
We separate the deduplication within a VM and cross VMs
and simplify the cross-VM deduplication while maximizing the inter-VM deduplication efficiency  as much as possible.
This is because global deduplication that detects the appearance of a chunk 
in any VM requires a substantial resource for fingerprint comparison.
To simplify cross-VM deduplication, we restrict the search scope of deduplication within the top $k$ 
most popular items. This popular data set is called the \textbf{PDS}. 
%This step accomplishes the canonical global fingerprint lookup using a popular fingerprint index.
 \begin{figure}
 \centering
  \includegraphics[width=3in]{figures/zipf_count_rank.pdf}
 \caption{Duplicate frequency versus  chunk ranking in a log scale after local deduplication.}
 \label{fig:Datazipf}
 \end{figure}

% {\it [Need to find a place to put these numbers in: Total number of chunks 
% in 350 snapshots: 1,546,635,485. 
% Total number of chunks after localized dedup: 283,121,924. Total number of unique chunks: 87,692,682.]}
What  motivates us to use PDS
is that for those chunks that appear in different VMs, the top $k$ most popular items
dominate the appearance distribution following our previous evaluation~\cite{WeiZhangIEEE}.  
Figure~\ref{fig:Datazipf} shows the distribution of chunk popularity for a data trace 
from Alibaba with 2500 VMs discussed in Section ~\ref{sect:evaluation}.
We define chunk popularity as the number of unique copies of the chunk in the data-store,
i.e., the number of copies of the chunk after local deduplication.
%the distribution of chunk popularity~\cite{WeiZhangIEEE} 
%based on a production data trace from Alibaba with over 2500 VMs.
The distribution from this figure is Zipf-like. 
Let  $\sigma$ denote as the percentage of unique data chunks belonging to PDS and 
from the evaluation in Section~\ref{sect:evaluation}, we find that
$\sigma$ with about  2\% can deliver a fairly competitive deduplication efficiency.
%$\sigma$ with about  range of 2 to 4\% can deliver a fairly competitive deduplication efficiency.
%For example,
%with top 2\% of the most popular chunks stored in PDS, up-to 40\% of the raw backup 
%data can be eliminated by fingerprint lookup in PDS index.

@@

\comments{
We have not adopted the index sampling in \cite{Guo2011} because its extension for
for a distributed architecture is difficult.  To use a distributed memory version of the sampled index, every deduplication 
request may access a remote machine for index lookup and the overall overhead of access latency for all requests
can be significant.
}

We describe two  complementary strategies below for the above  VM-centric design.
%duplicate search, 
%and %VM-centric file system block management. 
%\textbf{Cross-VM global deduplication using popular chunks} --
\begin{itemize}
\item 
\textbf{VM-specific duplicate search optimization} --
While cross-VM deduplication is simplified, we intend to optimize the VM-specific deduplication 
under a reasonable memory consumption
 to make up the loss of deduplication opportunities due to the cross-VM popularity constraint.
%Fingerprint lookup within a VM does not require a signficant  amount of memory resource.
We start with the standard version-based detection~\cite{Clements2009,Vrable2009}
to identify changed content with dirty bits in a coarse grain segment level.
%In our implementation 
%the segment size is 2MB
%and the device driver is extended to support tracking changed segments using a dirty bitmap. 
The reason to choose a coarse grain segment level is that 
since every write for a segment will touch a dirty bit, the device driver maintains dirty bits in 
memory and cannot afford a small segment size.
It should be noted that dirty bit tracking is supported or can be easily implemented in 
major virtualization solution vendors. 
%% this method is so widely-used and easy to understand, everyone include Pure Storage has it, I feel no need to explain further.
%{
%For example,
%the VMWare hypervisor has an API to let external backup applications know 
%the changed areas since last backup. 
%The Microsoft SDK provides an API that allows external applications to monitor 
%the VM's I/O traffic and implement such changed block tracking feature.
%}

@@

%The system allows all machines conduct backups in parallel, and each machine
%conducts the backup of one VM at a time, and thus only requires a write buffer for one VM.
\end{itemize}
The underlying distributed  file system has  a fixed replication degree for 
all file system blocks~\cite{googlefs03,hdfs10}.
We add extra replicas for file blocks containing PDS chunks  
and this brings additional  failure protection for these chunks shared by many VMs.
%Appendix~\ref{sect:fault} provides a detailed analysis of snapshot availability
%under this VC scheme and compares it with VO.
